---
title: "Process tracing"
subtitle: |
  | Metode 3
  | [.white[Link til slides: kortlink.dk/2acdv]](https://soerendamsbo.github.io/m3/m3-7_process-tracing.html)
author: Søren Damsbo-Svendsen<br>[.white[sdas@ifs.ku.dk]](mailto:sdas@ifs.ku.dk)
institute: Institut for Statskundskab<br>Københavns Universitet
date: "Uge 18"
output:
  xaringan::moon_reader:
    css: "style.css"
    self_contained: true
    nature:
      ratio: '16:9'
      highlightStyle: github
      highlightLines: true
      countincrementalSlides: false
---

```{r, include=F}
uge_nr <- 18

pacman::p_load(tidyverse, ggdag)
knitr::opts_chunk$set(echo = FALSE, fig.align = "center", fig.retina = 3, fig.height = knitr::opts_chunk$get("fig.width")/2)

pacman::p_load(tidyverse, lubridate, kableExtra)
Sys.setlocale(locale = "da_DK.UTF-8")
```


# Overblik

```{r plan}

start_date <- ymd("2021-02-11")
a_dates <- start_date+7*0:14
b_dates <- (start_date+1)+7*0:14

tibble(Uge = 6:20, 
       `Dato (holdtimer)` = tolower(paste0(day(a_dates),
                                           "-", 
                                           day(b_dates),
                                           " ", 
                                           format(a_dates, "%B %Y"))), 
       Emne = c("Kvantitativ indholdsanalyse",
                "Diskursanalyse I",
                "Diskursanalyse II",
                "interaktioner og modelspecifikation",
                "Logistisk regression",
                "Multilevel analyse",
                "Kausal inferens I: Designbaseret inferens",
                "Kausal inferens II: Paneldata",
                "Påskeferie",
                "Kausal inferens III: Eksperimentelle designs",
                "Kausal inferens IIII: Regression Discontinuity (RD)",
                "Social Data Science: Big Data",
                "Process tracing",
                "Kriterier for god videnskab",
                "Opsamling/spørgetime"), 
       `Øvelsesopgave` = c(1:8, "", 9:14)) %>%
  mutate(`Dato (holdtimer)` = ifelse(Uge %in% c(6,7,12,14,16,17,19,20), "", `Dato (holdtimer)`)) %>% 
  kable("html", 
        align = c("c", "c", "l", "c"),
        escape = F,
        table.attr = "style=\"line-height: 100%;\"") %>% 
  kable_styling(font_size = 16,
                bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>% 
  row_spec(9, bold = F, color = "grey") %>% 
  row_spec(uge_nr-5, bold = T, color = "white", background = "#8b2325")
```

---

# Recap fra sidste gang

---

# Dagens program

---

# Dagens formål

---

# Pensum

Beach, Derek og Rasmus Brun Pedersen. (2012). Process tracing: metode, design og forskningslogik. I: Lotte Bøgh Andersen, Kasper Møller Hansen og Robert Klemmensen (red.): *Metoder i statskundskab* (2. udgave, s. 235-257). København: Hans Reitzels Forlag.

Bennett, Andrew. (2008). Process Tracing: a Bayesian Perspective. I: Janet M. Box-Steﬀensmeier, Henry Brady og David Collier (red.): *The Oxford Handbook of Political Methodology* (s. 702-721). New York: Oxford University Press.

---


---
---

# Dagens pointer

---

# Næste gang

---

# Tak for i dag!
